---
title: "Examining the media trends of Rohingya Crisis 2010-2021"
output: github_document
editor_options: 
  chunk_output_type: console
author: Michelle Lee
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(readr)
library(broom)
library(dbplyr)
library(viridis)
library(reshape2)
library(plotly)
library(lubridate)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```
### Project Title

### About the Data: 
We searched for "Rohingya" on Google News from 01/01/2010 to 01/01/2021 to gather news articles that were published during the past decade. Using Parsehub, which is ... , I collected the title, publisher, publication date, the url, and a short description of each news article.  

``` {r importing data}
news_data = read_csv (file = "./rohingya_news_data_20200104.csv",
                     col_types = cols(
                       Date = col_date(format = "%mm/%dd/%yy"),
                       `Subways: % Change From 2019 Equivalent Day` = col_number(),
                       `Buses: % Change From 2019 Equivalent Day` = col_number(),
                       `Bridges and Tunnels: % Change From 2019 Equivalent Day` = col_number()
                       ) #only changed the formats of important variables 
) %>%
  janitor::clean_names()
skimr::skim(mta_data)
mta_data =
  mta_data %>%
  subset(select = -c(lirr_total_estimated_ridership, lirr_percent_change_from_2019_monthly_weekday_saturday_sunday_average, metro_north_total_estimated_ridership, metro_north_percent_change_from_2019_monthly_weekday_saturday_sunday_average, access_a_ride_total_scheduled_trips, access_a_ride_percent_change_from_2019_monthly_weekday_saturday_sunday_average, bridges_and_tunnels_total_traffic, bridges_and_tunnels_percent_change_from_2019_equivalent_day))
#exclude data for lirr, metronorth, access-a-ride, bridges & tunnel
```
Now we will calculate the actual 2019 ridership based on the percent change data

``` {r calculate 2019 ridership bus subway}
mta_data = mta_data %>%
  mutate( 
    'subway_2019' = subways_total_estimated_ridership/(1+(subways_percent_change_from_2019_equivalent_day/100)),
    'bus_2019'=
      buses_total_estimated_ridership/(1+(buses_percent_change_from_2019_equivalent_day/100))
    ) %>%
  rename(
    "subway_2020" = subways_total_estimated_ridership,
    "subway_pct_change" = subways_percent_change_from_2019_equivalent_day,
    "bus_2020" = buses_total_estimated_ridership,
    "bus_pct_change" = buses_percent_change_from_2019_equivalent_day
    )
```

```{r plotting general trends of MTA ridership 2019 & 2020}
#change date to date format and order by date
plot_mta =
  mta_data %>%
  mutate(
    date= as.Date(date,format = "%m/%d")) %>%
  arrange(date)
#create a text_label label
text_label =
  plot_mta %>%
  mutate(text_label = str_c("percent change from 2019 to 2020: ", subway_pct_change)) %>%
  select(date, text_label)

#graph for subway 
#pivoting
plot_subway = 
  plot_mta %>%
  select(subway_2020, subway_2019, date) %>%
  melt(., id.vars = "date") %>%
#merge based on month 
  merge(text_label, by = "date") %>%
  #plotting
  plot_ly(
    x = ~date, y = ~value, type = "scatter", mode = "markers",
    color = ~variable, text = ~text_label) %>%
  layout (
    title = "Subway Ridership Trends 2019 - 2020",
    xaxis = list(title ="Month/Day", tickformat = "%m/%d"), #drop year
    yaxis = list(title="Ridership"))  %>%
  add_lines(x =as.Date("2020-03-01"), line = list(dash="dot", color = 'red', width=0.5, opacity = 0.5),name = 'First case on 3/1') %>%
  add_lines(x =as.Date("2020-04-07"), line = list(dash="dot", color = 'red', width=0.5, alpha = 0.5),name = '100K cases in NYC on 04/07') %>%
  add_lines(x =as.Date("2020-05-26"), line = list(dash="dot", color = 'red', width=0.5, alpha = 0.5),name = '200K cases in NYC on 05/26')
#graph for bus

#pivoting
plot_bus = 
  plot_mta %>%
  select(bus_2020,bus_2019,date) %>%
  melt(., id.vars = "date") %>% 
#merge based on month 
  merge(text_label, by = "date") %>%
#plotting
  plot_ly(
    x = ~date, y = ~value, type = "scatter", mode = "markers",
    color = ~variable, text = ~text_label) %>%
  layout (
    title = "Bus Ridership Trends 2019 - 2020",
    xaxis = list(title ="Month/Day", tickformat = "%m/%d"),
    yaxis = list(title="Ridership")) %>%
  add_lines(x =as.Date("2020-03-01"), line = list(dash="dot", color = 'red', width=0.5, opacity = 0.5),name = 'First case on 3/1') %>%
  add_lines(x =as.Date("2020-04-07"), line = list(dash="dot", color = 'red', width=0.5, alpha = 0.5),name = '100K cases in NYC on 04/07') %>%
  add_lines(x =as.Date("2020-05-26"), line = list(dash="dot", color = 'red', width=0.5, alpha = 0.5),name = '200K cases in NYC on 05/26')

#ggplot for subway
plot_subway_2 = 
  plot_mta %>%
  select(subway_2020, subway_2019, date) %>%
  melt(., id.vars = "date") %>%
  ggplot(aes(x=date, y=value, color=variable)) +
  geom_point(alpha=0.5) +
  geom_smooth(se = FALSE)  +
  xlab("Month/Day") +
  ylab("Ridership") +
  ggtitle("Subway Ridership Trends 2019 - 2020") +
  geom_vline(xintercept=as.Date("2020-03-01"), linetype="dotted", color = 'red')+
  geom_vline(xintercept=as.Date("2020-04-07"), linetype="dotted", color = 'red')+
  geom_vline(xintercept=as.Date("2020-05-26"), linetype="dotted", color = 'red')+
  geom_text(x=as.Date("2020-03-01"), y=6100000, label ="First case", angle=90, vjust = 1.2, size=3,color='black') +
  geom_text(x=as.Date("2020-04-07"), y=6100000, label ='100K cases', angle=90, vjust = 1.2, size=3,color='black')+
  geom_text(x=as.Date("2020-05-26"), y=6100000, label ='200K cases', angle=90, vjust = 1.2, size=3,color='black')

#ggplot for bus
plot_bus_2 = 
  plot_mta %>%
  select(bus_2020, bus_2019,date) %>%
  melt(., id.vars = "date") %>% 
  ggplot(aes(x=date, y=value, color=variable)) +
  geom_point(alpha=0.5) +
  geom_smooth(se = FALSE) +  
  xlab("Month/Day") +
  ylab("Ridership") +
  ggtitle("Bus Ridership Trends 2019 - 2020") +
  geom_vline(xintercept=as.Date("2020-03-01"), linetype="dotted", color = 'red')+
  geom_vline(xintercept=as.Date("2020-04-07"), linetype="dotted", color = 'red')+
  geom_vline(xintercept=as.Date("2020-05-26"), linetype="dotted", color = 'red')+
  geom_text(x=as.Date("2020-03-01"), y=2400000, label ="First case", angle=90, vjust = 1.2, size=3,color='black') +
  geom_text(x=as.Date("2020-04-07"), y=2400000, label ='100K cases', angle=90, vjust = 1.2, size=3,color='black')+
  geom_text(x=as.Date("2020-05-26"), y=2400000, label ='200K cases', angle=90, vjust = 1.2, size=3,color='black')

#see the two graphs together:
plot_bus
plot_subway
(plot_bus_2+plot_subway_2)
```
#source for the vertical lines:
https://edition.cnn.com/2020/03/02/us/new-york-coronavirus-first-case/index.html (3/1)
https://www1.nyc.gov/site/doh/covid/covid-19-data.page

Let us first look at the average ridership of subway during 2019 and 2020 for each month
```{r}
#separate by month and day
mta_data = 
  mta_data %>%
  arrange(date) %>%
  separate(date, into = c("month", "day", "year"))%>%
  mutate(month = as.numeric(month),
         day = as.numeric(day)) %>%
  select(-c(year)) #drop year column

mta_subway_ridership =
  mta_data %>%
  group_by(month)%>%
  summarize(
    avg_subway_2019 = mean(subway_2019),
    avg_subway_2020 = mean(subway_2020)
  )
mta_subway_ridership%>%knitr::kable()
```
Let us now compare 2019 and 2020 ridership by month and check if they are significantly different.

First we need to create our dataframe that nests the observations for 2019 and 2020 subway ridership for each month.
```{r}
mta_2019_sample = 
  mta_data%>%
  select(month, subway_2019) %>%
  nest(subway_2019)%>%
  mutate("subway_2019_sample" = data)%>%
  select(-data)

mta_2020_sample = 
  mta_data%>%
  select(month, subway_2020)%>%
  nest(subway_2020)%>%
  mutate("subway_2020_sample" = data)%>%
  select(-data)

mta_samples = 
  bind_cols(mta_2019_sample, mta_2020_sample)%>%
  select(-month...3)%>%
  rename(month = month...1)
```

Now we need to perform a t-test to check if the subway ridership is signficantly different between 2019 and 2020 data. We map across each month to test if the values of the samples of 2019 and 2020 ridership data differs for each month. We then compare the p-value obtained from the tests to determine whether the differene in ridership numbers is significant or not

```{r}
mta_t_test = 
  mta_samples%>%
  mutate(t_test = map2(.x = subway_2019_sample, .y = subway_2020_sample, ~t.test(.x , .y) ),
         t_test_results = map(t_test, broom::tidy))%>%
  select(month, t_test_results)%>%
  unnest(t_test_results)%>%
  select(month,p.value)%>%
  mutate(difference = case_when(
    p.value >= 0.05 ~ "insignificant",
    p.value < 0.05 ~ "significant"),
    p.value = ifelse(
    p.value< 0.001,"<0.001",round(p.value, digits = 4))) %>%
  arrange(month) 

knitr::kable(mta_t_test, digits = 3)
```
We can merge the t-test results with the average riderships from before..

```{r}
mta_year_ttest = 
  bind_cols(mta_subway_ridership, mta_t_test)%>%
  select(-month...4)%>%
  rename(month = month...1) 
knitr::kable(mta_year_ttest)
  
```

Using this final dataframe, we can plot a graph showing the average ridership for each month comparing the 2019 to 2020 ridership and then labelling the signficance or lack of significance in ridership differences..

[Plot needs to be created]
``` {r plotting}
#create a text_label label
text_label =
  mta_year_ttest %>%
  mutate(text_label = str_c("p-value: ",p.value, "\nDifference: ", difference)) %>%
  select(month, text_label)
#pivoting
plot_ttest = 
  mta_subway_ridership %>%
  rename(
    "2020"=avg_subway_2020,
    "2019"=avg_subway_2019
  ) %>%
  melt(., id.vars = "month") %>%
#merge based on month 
  merge(text_label, by = "month")

#plotting
plot_ttest %>%
plot_ly(
    x = ~month, y = ~value, type = "scatter", mode = "lines+markers",
    color = ~variable, text = ~text_label) %>%
  layout (
    title = "Monthly Average Ridership of Subway 2019 vs 2020",
    xaxis = list(title ="Months",range=c(3,11)),
    yaxis = list(title="Average Ridership"))
```

Let us now compare subway and bus ridership in 2020 by month and check if they are significantly different.

First we need to propose a regression model for subway ridership in 2020.
As the outcome (subway ridership) is continuous, we can fit a linear regression model. 

```{r}
subway_ridership = lm(subway_2020 ~ month, data = mta_data)
subway_ridership %>% 
  broom::tidy() 

#to tidy the output and get only the intercept, slope and p-values 
subway_ridership %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

#to plot for regression line 
  ggplot(mta_data, aes(month, subway_2020)) +
  geom_point() +
  stat_smooth(method = lm)
```

* Show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
mta_data %>% 
  modelr::add_predictions(subway_ridership) %>% 
  modelr::add_residuals(subway_ridership) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point() + 
  labs(x = "Predicted value", 
       y = "Residual")
```

Next, we need to propose a regression model for bus ridership in 2020.
As the outcome (bus ridership) is continuous, we can fit a linear regression model. 

```{r}
bus_ridership = lm(bus_2020 ~ month, data = mta_data)
bus_ridership %>% 
  broom::tidy() 

#Now we need to tidy the output and get only the intercept, slope and p-values 
bus_ridership %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

#to plot for regression line
ggplot(mta_data, aes(month, bus_2020)) +
  geom_point() +
  stat_smooth(method = lm)
```
